A``Periodic Table'' for Bravais Lattices

### 2024-08-03 DeloneGrid/DeloneGrid.tex
HorizontalDeloneGrid -- updaed Delone Grid paper - started

%xxxxxx just for testing github % 2
%xxxxxx just for testing github % 3

\documentclass[preprint]{iucr}              % DO NOT DELETE THIS LINE
\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
%\usepackage{bm}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{booktabs}
%\usepackage{calligra}
\usepackage{array}
\DeclareMathAlphabet{\mathcalligra}{T1}{calligra}{m}{n}
\def\mathbi#1{\textbf{\em #1}}
\numberwithin{equation}{section}
%\DeclareMathSymbol{\Gamma}{\mathalpha}{letters}{"00}
%\DeclareMathSymbol{\Lambda}{\mathalpha}{letters}{"03}
%\DeclareMathSymbol{\Omega}{\mathalpha}{letters}{"0A}
%\DeclareMathAlphabet{\mathitbf}{OML}{cmm}{b}{it}
\hyphenation{Niggli}
\def\mathbi#1{\textbf{\em #1}}
%\numberwithin{equation}{section}
%\DeclareMathSymbol{\Gamma}{\mathalpha}{letters}{"00}
%\DeclareMathSymbol{\Lambda}{\mathalpha}{letters}{"03}
%\DeclareMathSymbol{\Omega}{\mathalpha}{letters}{"0A}
%\DeclareMathAlphabet{\mathitbf}{OML}{cmm}{b}{it}
\usepackage{color}
\usepackage{ulem}
\usepackage{url}
\usepackage{yfonts}
%\usepackage{xr-hyper}
%\usepackage[draft]{hyperref}
%\usepackage{bibentry}
\newcommand{\SVI}[0]{$\bf{S^{6}}$}
\newcommand{\GVI}[0]{$\bf{G^{6}}$}
\newcommand{\CIII}[0]{$\bf{C^{3}}$}
\newcommand{\DVII}[0]{$\bf{D^{7}}$}
\newcommand{\VVII}[0]{$\bf{V^{7}}$}

@@

%-------------------------------------------------------------------------
% Information about journal to which submitted
%-------------------------------------------------------------------------
\journalcode{A}              % Indicate the journal to which submitted
%   A - Acta Crystallographica Section A
%   B - Acta Crystallographica Section B
%   C - Acta Crystallographica Section C
%   D - Acta Crystallographica Section D
%   E - Acta Crystallographica Section E
%   F - Acta Crystallographica Section F
%   J - Journal of Applied Crystallography
%   M - IUCrJ
%   S - Journal of Synchrotron Radiation
\makeatletter
\font\dummyft@=dummy \relax
\makeatother

@@

	% The title of the paper. Use \shorttitle to indicate an abbreviated title
	% for use in running heads (you will need to uncomment it).
	{\LARGE \emph{\today}} \\
	\title{Delone lattice studies in \CIII, the space of three complex variables}
	%\title{Note on the transformation of three-space basis vectors to  corresponding matrix for Delaunay scalars}
	\shorttitle{properties of C3}

@@

	\aff[a]{Ronin Institute, 9515 NE 137th St, Kirkland, WA, 98034-1820 \country{USA}}
	\aff[b]{Ronin Institute, c/o NSLS-II, Brookhaven National Laboratory, Upton, NY, 11973 \country{USA}}

@@

	\keyword{lattice}
	\keyword{reduction}
	\keyword{Delone}
	\keyword{Selling}
	\keyword{\CIII}

@@

	\begin{synopsis}
		The space \CIII{} is explained in more detail than
		in the original description. Boundary transformations
		of the fundamental unit are described in detail. 
		A graphical presentation of the basic coordinates
		is described and illustrated.
	\end{synopsis}
	\newcommand{\si}[0]{$s_1$}
	\newcommand{\sii}[0]{$s_2$}
	\newcommand{\siii}[0]{$s_3$}
	\newcommand{\siv}[0]{$s_4$}
	\newcommand{\sv}[0]{$s_5$}
	\newcommand{\svi}[0]{$s_6$}
	\newcommand{\Svec} [0] {\{\si, \sii, \siii, \siv, \sv, \svi \}}
	\newcommand{\SvecA} [0] {\{-\si, -\si+\sii, \si+\siii, \si+\sv, \si+\siv, \si+\svi \}}

@@

	\begin{abstract}
		The Delone (Selling) scalars, which are used in 
		unit cell reduction and in lattice type determination,
		are studied in \CIII, the space of three complex variables.
		The three complex coordinate planes are composed of the six
		Delone scalars. The transformations at boundaries of the
		Selling reduced orthant are described as matrices of operators.
		A graphical representation as the projections onto the
		three coordinates is described.

@@

	\section{Introduction}

@@

	~\citeasnoun{andrews2019b} chose to 
	represent the Selling scalars in the space \SVI{},
	\Svec{} (defined in the order above), 
	as a way to create a metric space
	for the measurement of the distance between lattices. 
	They also considered the representation 
	of this space as the
	space of three complex dimensions, \CIII{} or 
	{$\{c_1$}, {$c_2$}, {$c_3\}$}.

	In \CIII{}, in terms of the Selling scalars, 
	a vector is defined as \{(\si,\siv ), (\sii,\sv),(\siii,\svi)\}, 
	where the real and imaginary parts
	of each are the ``opposite'' scalars 
	according to the definition of~\citeasnoun{Delaunay1932}
	(opposite in terms of the Bravais tetrahedron of scalars) (see~\citeasnoun{andrews2019a}).
	As a mnemonic device, 
	note that the complex components involve ($\alpha,a$), ($\beta, b$), and ($\gamma,c$).
	Additionally, each complex term uses all 
	four three-space vectors; for example, $c_1$ is ($b \cdot c$, $a \cdot d$).

	~\citeasnoun{andrews2019b} considered the matrix representations 
	of the reflections in \CIII{} (and in \SVI{}). 
	This paper describes the boundary transformations 
	at the edges of the fundamental	unit of \CIII{}
	and also a graphical presentation. 
	In \SVI{}, the fundamental unit is the all negative orthant, 
	which contains only, and all of, the reduced cells. 
	In \SVI{} and \CIII{}, 
	the boundaries located where any \SVI{} scalar 
	(or correspondingly in \CIII{}, a real or imaginary part) 
	equals zero.

	\section{Notation}

	Complex numbers are represented 
	in Cartesian format $(x,y)$, 
	where $x$ is the real part and
	$y$ is the imaginary part.

	We represent a vector 
	in \CIII{} by \{$(x_1,y_1)$, $(x_2,y_2)$, $(x_3, y_3)$\} 
	as an alternative to  \{(\si,\siv ), (\sii,\sv),(\siii,\svi)\}.

	Next we define the operators in \CIII{} that 
	will be used in the matrix descriptions of 
	the transformations at the boundaries of the fundamental unit,
	see Table ~\ref{tab:operators}.

	\begin{table}	
		\label{tab:operators}
		\caption{The Result column values give the products of each operator as applied to $(x_j, y_j)$}.
		\begin{tabular}{c l l l }
			\toprule
			Operator			&Usage						&Result					&Name\\
			\midrule
			$\Minus{}_r $		&  $\Minus{}_r(c_j)$ 		& $(-x_j, -x_j+y_j)$ 	& Minus real \\ 
			$\Minus{}_i $ 		& $\Minus{}_i(c_j)$ 		& $(x_j-y_j,-y_j) $ 	& Minus imag\\ 
			$\mathcal{P}_r$ 	& $ \mathcal{P}_r(c_j)$ 	& $(x_j, x_j)$ 		& Plus real\\ 
			$\mathcal{P}_i$ 	&  $\mathcal{P}_i(c_j)$  	& $(y_j, y_j)$ 		& Plus imag\\ 
			$\Real$ 			&  $\Real(c_j)$ 			& $x_j$ or $(x_j,0)$		& Real \\ 
			$\Imaginary$ 		&  $\Imaginary(c_j)$ 		& $y_j$ or $(y_j,0)$ 		& Imaginary\\
			\bottomrule
		\end{tabular}		
	\end{table}

	\section{Matrices of boundary transformations}
	~~\\

	\citeasnoun{andrews2019b} listed the boundary transformations
	for each of the six \SVI{} boundaries. For each,
	they only listed 2 transformations, following \citeasnoun{Delaunay1932}
	and \citeasnoun{Delone1975}. That list is incomplete in the sense
	that applying the 24 reflections gives 24 boundary transformation,
	four for each boundary, rather than 12.

	The C3 vector transformations are represented as multiplication by a matrix of the operators defined in Table 1. When a complex vector component is 'multiplied' by an operator, the result is as defined in the table."

	For the boundary at $s_1$: (the real component of $c_1$).

	$\begin{bmatrix}
		\Minus{}_r		& 0					& 0 \\
		\Plus{}_r		&  \textrm{i}\Real{}	& \Real{} \\
		\mathcal{P}_r	& \textrm{i}\Imaginary{}	& \Imaginary{}
	\end{bmatrix}$
	%
	$\begin{bmatrix}
		\Minus{}_r	& 0 				& 0 \\
		\Plus{}_r	&\textrm{i}\Imaginary{}	& \Imaginary{} \\
		\Plus{}_r	& \textrm{i}\Real{}	& \Real{} \\
	\end{bmatrix}$ 
	%
	$\begin{bmatrix}
		\label{thirdmatrix}
		\Minus{}_r	& 0			& 0 \\
		\Plus{}_r	&  \Real{}	& \textrm{i}\Real{} \\
		\Plus{}_r	& \Imaginary{}		&\textrm{i}\Imaginary{}
	\end{bmatrix}	$
	%
	$\begin{bmatrix}
		\Minus{}_r	& 0			& 0 \\
		\Plus{}_r	& \Imaginary{}		&\textrm{i}\Imaginary{} \\
		\Plus{}_r	&  \Real{}	&\textrm{i}\Real{} \\
	\end{bmatrix}$ \\

	For the boundary at $s_4$: (the imaginary component of $c_1$).

	$\begin{bmatrix}
		\Minus{}_i	& 0					& 0 \\
		\Plus{}_i	&  \textrm{i}\Real{}	& \Real{} \\
		\Plus{}_i	& \textrm{i}\Imaginary{}	& \Imaginary{}
	\end{bmatrix}$
	%
	$\begin{bmatrix}
		\Minus{}_i	& 0					& 0 \\
		\Plus{}_i	& \textrm{i}\Imaginary{}	& \Imaginary{} \\
		\Plus{}_i	&  \textrm{i}\Real{}	& \Real{} \\
	\end{bmatrix}$ 
	%
	$\begin{bmatrix}
		\Minus{}_i	& 0			& 0 \\
		\Plus{}_i	&  \Real{}	& \textrm{i}\Real{} \\
		\Plus{}_i	& \Imaginary{}		& \textrm{i}\Imaginary{}
	\end{bmatrix}	$
	%
	$\begin{bmatrix}
		\Minus{}_i	& 0		& 0 \\
		\Plus{}_i	& \Imaginary{}	& \textrm{i}\Imaginary{} \\
		\Plus{}_i	&  \Real{}& \textrm{i}\Real{} \\
	\end{bmatrix}$ \\

	For the boundary at \sii{} (the real component of \cii{}):

	$\begin{bmatrix}

		\textrm{i}\Real{}	& \Plus{}_r		& \Real{} \\
		0				& \Minus{}_r	& 0 \\
		\textrm{i}\Imaginary{}	& \Plus{}_r		& \Imaginary{}
	\end{bmatrix}$
	%
	$\begin{bmatrix}
		\textrm{i}\Imaginary{}	& \Plus{}_r		&  \Imaginary{} \\
		0				& \Minus{}_r	& 0 \\
		\textrm{i}\Real{}	& \Plus{}_r		& \Real{} \\
	\end{bmatrix}$
	%
	$\begin{bmatrix}
		\Real{}	&\Plus{}_r		&   \textrm{i}\Real{} \\
		0		& \Minus{}_r	&  0 \\
		\Imaginary{}	&\Plus{}_r		& \textrm{i}\Imaginary{}
	\end{bmatrix}	$ 
	%
	$\begin{bmatrix}
		\Imaginary{}	& \Plus{}_r	& \textrm{i}\Imaginary{} \\
		0		&\Minus{}_r	&  0 \\
		\Real{}	& \Plus{}_r	&  \textrm{i}\Real{} \\
	\end{bmatrix}$ \\

	For the boundary at \sv: (the imaginary component of \cii{}).

	$\begin{bmatrix}
		\textrm{i}\Real{}	& \Plus{}_i		& \Real{} \\
		0				& \Minus{}_i	& 0 \\
		\textrm{i}\Imaginary{}	& \Plus{}_i		& \Imaginary{}
	\end{bmatrix}$
	%
	$\begin{bmatrix}
		\textrm{i}\Imaginary{}	& \Plus{}_i		&  \Imaginary{} \\
		0				& \Minus{}_i	& 0 \\
		\textrm{i}\Real{}	& \Plus{}_i	& \Real{} \\
	\end{bmatrix}$ 
	%
	$\begin{bmatrix}
		\Real{}	&\Plus{}_i		&   \textrm{i}\Real{} \\
		0		& \Minus{}_i	&  0 \\
		\Imaginary{}	&\Plus{}_i		& \textrm{i}\Imaginary{}
	\end{bmatrix}	$ 
	%
	$\begin{bmatrix}
		\Imaginary{}	& \Plus{}_i	& \textrm{i}\Imaginary{} \\
		0		&\Minus{}_i	&  0 \\
		\Real{}	& \Plus{}_i	&  \textrm{i}\Real{} \\
	\end{bmatrix}$ \\

	For the boundary at \siii{} (the real component of \ciii{}):

	$\begin{bmatrix}

		\textrm{i}\Real{}	& \Real{}	& \Plus{}_r \\
		\textrm{i}\Imaginary{}	& \Imaginary{}	& \Plus{}_r  \\
		0				& 0		& \Minus{}_r  \\
	\end{bmatrix}$
	%
	$\begin{bmatrix}
		\textrm{i}\Imaginary{}	&  \Imaginary{}	& \Plus{}_r \\
		\textrm{i}\Real{}	& \Real{}		& \Plus{}_r \\
		0				& 0			& \Minus{}_r \\
	\end{bmatrix}$ 
	%
	$\begin{bmatrix}
		\Real{}	&   \textrm{i}\Real{}	&\Plus{}_r \\
		\Imaginary{}	& \textrm{i}\Imaginary{}	&\Plus{}_r \\
		0 		&  0				& \Minus{}_r \\
	\end{bmatrix}	$ 
	%
	$\begin{bmatrix}
		\Imaginary{}	& \textrm{i}\Imaginary{}	& \Plus{}_r \\
		\Real{}	&  \textrm{i}\Real{}	& \Plus{}_r \\
		0		&  0				&\Minus{}_r \\
	\end{bmatrix}$ \\

	For the boundary at \svi{} (the imaginary component of \ciii{}):

	$\begin{bmatrix}

		\textrm{i}\Real{}	& \Real{}	& \Plus{}_i \\
		\textrm{i}\Imaginary{}	& \Imaginary{}	& \Plus{}_i \\
		0				& 0		& \Minus{}_i 
	\end{bmatrix}$
	%
	$\begin{bmatrix}
		\textrm{i}\Imaginary{}	&  \Imaginary{}	& \Plus{}_i \\
		\textrm{i}\Real{}	& \Real{}		& \Plus{}_i \\
		0				& 0			& \Minus{}_i \\
	\end{bmatrix}$ 
	%
	$\begin{bmatrix}
		\Real{}	&   \textrm{i}\Real{} &\Plus{}_i \\
		\Imaginary{}	& \textrm{i}\Imaginary{}	&\Plus{}_i \\
		0		&  0				& \Minus{}_i \\
	\end{bmatrix}	$ 
	%
	$\begin{bmatrix}
		\Imaginary{}	& \textrm{i}\Imaginary{}	& \Plus{}_i \\
		\Real{}	&  \textrm{i}\Real{}	& \Plus{}_i \\
		0		&  0				&\Minus{}_i 
	\end{bmatrix}$ \\

\subsection{Derivation}

Consider the \SVI{} vector 	
	\{$(x_1,y_1)$, $(x_2,y_2)$, $(x_3, y_3)$\};
in \CIII{} that vector is:  \{(\si,\siv ), (\sii,\sv),(\siii,\svi)\}.
Stated in \CIII{}, the problem is to find a matrix that converts a \CIII{}
vector to a more reduced one. Take the case of \si positive and the
other scalars negative.\\

 \{(-\si,~\siv-\si ),~(\siii+\si,~\sii+\si), (\svi+\si,~\sv+\si)\}\\
 =
\begin{tabular}{ccc}

 	RedMat

$ \begin{Bmatrix}
 $(\si,\siv )$\\ $(\sii,\sv)$\\$(\siii,\svi)$
\end{Bmatrix}$
\end{tabular}\\

We need to solve for RedMat.
Here, one of the four possible result vectors has been chosen 
(arbitrarily).

Element 1 of result vector is (-\si,~\siv-\si ), and
it is created by multiplying the beginning \CIII{} vector by the
first row of the transformation matrix. In this case, we see
that \ci of the result only contains scalars from \ci of
the input, so the first row must contain some operator to
be multiplied times input \ci, and the second and third
elements of the matrix row must be zero. We create an operator
($Minus~real$) to transform the input \ci to the output \ci; so
the first row of the reduction matrix is: $\Minus{}_r	 0		 0$.

Element 2 of the result vector is (\siii+\si,~\sii+\si),
and
it is created by multiplying the beginning \CIII{} vector by the
second row of the transformation matrix. The result \cii
contains \si in both the real and imaginary parts. So
the first matrix element in that row must insert \si, the
real part of the input \ci, into both the real and imaginary
parts. An operator to produce that is created 
(Plus{}~real). Next we see that the result contains \sii and \siii.
\siii appears in the real part of the result element and in the
third part of the input vector. So the third element of the 
second matrix row must extract the \siii, the real part of 
\ciii of the input. So the second row now contains the Plus real
operator and the element to put the  real part of \ciii of the input.
The final matrix element of the second row (the second element)
must extract the real part of input \cii and place it in the
imaginary part of the output vector. The resulting second
row of the matrix is $\Plus_r~i\Real~\Real$.

Using similar logic, the third row of the matrix is:
$\Plus_r~i\Imaginary~\Imaginary$.

Having derived the first matrix, the other 23 can
be derived by the same process. However because all the
\SVI boundaries are related by the 24 reflection, the
other 23 matrices can be generated simply
by multiplying by each of the reflection matrices.

\subsection{Example C++ implementation}

\begin{verbatim}
static std::complex<double> c3_Minus_Real(std::complex<double>& a) {
   const std::complex<double> c = { -a.real(), -a.real() + a.imag() };
   return c;
}

C3 C3::c3_s1(const C3& a) {
   C3 c;
   c[0] = c3_Minus_Real(a[0]);
   c[1] = c3_Plus_Real(a[0]) + c3_Real(a[1]) +
   c3_I_Times_Real(a[2]);
   c[2] = c3_Plus_Real(a[0]) + c3_Imag(a[1]) +
   c3_I_Times_Imag(a[2]);
   return c;
}
\end{verbatim}

\subsection{Example of use}

To demonstrate the application of the above matrices,
example C++ code was written to create matrices of operators;
see the supplementary material for the code or the complete
library reference in the Availability of Code below.

Create a \CIII{} vector c3 with
the values \{(1,2),(3,4),(5,6)\} (which does not
correspond to a valid unit cell). Multiply it by the matrix
of operators derived above.

The matrix multiplication can be implemented as follows (including
only one example of the functions for operators (c3\_Minus\_Real))"

Therefore, RedMat = \\
\begin{center}
	$\begin{bmatrix}
		\Minus{}_r	& 0			& 0 \\
		\Plus{}_r	&  \textrm{i}\Real{}	& \Real{} \\
		\Plus{}_r	& \textrm{i}\Imaginary{}		&\Imaginary{}
	\end{bmatrix}	$
\end{center}

For example:

\begin{center}
	\{(-1.000,1.000) (4.000,6.000) (5.000,7.000)\} =\\
	$\begin{bmatrix}
		\Minus{}_r	& 0			& 0 \\
\Plus{}_r	&  \textrm{i}\Real{}	& \Real{} \\
\Plus{}_r	& \textrm{i}\Imaginary{}		&\Imaginary{}
	\end{bmatrix}	$
	$\begin{Bmatrix}
		(1,2)\\
		(3,4)\\
		(5,6)
	\end{Bmatrix}$\\

\end{center}

which can be verified by hand.

	\section{Basics}

	The standard representation of the identity operation is\\

	$\bf{c'}$  = 
	$\begin{bmatrix}
		$1$	& $0$	&  $0$ \\
		$0$	& $1$	& $0$ \\
		$0$	& $0$	& $1$ \\
	\end{bmatrix}$ 
	$\bf{c}$. \\

	The identity in \CIII{} could also be written:\\

	$\bf{c'}$  = 
	$\begin{bmatrix}
		$1$	& $0$						&  $0$ \\
		$0$	& \Real{} +\textrm{i}\Imaginary{}	& $0$ \\
		$0$	& $0$						& \Real{} +\textrm{i}\Imaginary{} \\
	\end{bmatrix}$ 
	$\bf{c}$.
	\\
	which is simply replacement of the value one with an
	equivalent representation as operators.

	~\citeasnoun{Delaunay1932} does not 
	consider the boundary transformations 
	in detail. 
	However, he uses them to 
	define the process of Selling reduction. 
	For example in \SVI{}, 
	he lists the following as one of the possible results 
	for a transformation on \si{}, translated to \SVI{}:
	\SvecA{}.
	The third boundary transform for $s_1$ above
	implements this operation and interchanges the
	real part of \ciii{} and the imaginary part of \cii{}:

	\begin{center}
		$\begin{bmatrix}
			\Minus{}_r	& 0			& 0 \\
			\Plus{}_r	&  \Real{}	& \textrm{i}\Real{} \\
			\Plus{}_r	& \Imaginary{}		&\textrm{i}\Imaginary{}
		\end{bmatrix}	$
	\end{center}

	Considered in \CIII, Delone's alternate transformation 
	for the \si{} boundary would
	exchange the real of \cii{} 
	with the imaginary part of \ciii{}. That is the fourth
	matrix above in the list for \si{}. 
	The other two transformations for \si{} can 
	be generated from the two we have 
	just mentioned by the ``exchange operation''~\cite{andrews2019b} 
	applied to the second and third \CIII{} coordinates. Delone 
	did not describe the latter two transformations, 
	perhaps because even a single transformation 
	was adequate to implement reduction. He had already listed two.

	\section{Graphical display of projections}

	The two-dimensional nature of the three coordinates of \CIII{} 
	suggests their use for graphical display.

	As an example, we use Phospholipase A2 (\textbf{PLA2}) (retrieved from 
	the Protein Data Bank \cite{Bernstein1977}), which has had several similar
	or identical structures determined \cite{LeTrong2007}. \citeasnoun{andrews2019b}
	found additional cases (see Table \ref{PLA2})

	\begin{table}
		\begin{tabular}{m{0.08\textwidth} c c c c c c c c c}
			\toprule
			PDB id & Centering & a&b&c&$\alpha$&$\beta$&$\gamma$\\
			\midrule
			1DPY & hR& 57.98& 57.98& 57.98& 92.02& 92.02& 92.02\\
			~~\& 1FE5 \\
			1G0Z  & hP& 80.36& 80.36& 99.44& 90&    90&    120\\
			~~\& 1U4J\\
			1G2X & mC& 80.95& 80.57& 57.1 & 90&    90.35&  90\\
			2OSN & hR& 57.10& 57.10& 57.10& 89.75& 89.75&  89.75\\
			\bottomrule
		\end{tabular}	\\
		\caption{Phospholipase A2 unit cells}
		\label{PLA2}
	\end{table}

	Below, Figure \ref{fig1} shows the unit cells as 
	reported (the centering of lattices has not been removed).
	The following figures (Figures \ref{fig2} to \ref{fig5}) show various transformations and
	embellishments of the reported cells.

	\begin{figure}
		\includegraphics[width=1.0\textwidth]{1_35}
		\caption{Phospholipase A2 unit cells as reported. }
		\label{fig1}
	\end{figure}

	\begin{figure}
		\includegraphics[width=1.0\textwidth]{2_06}
		\caption{The PLA2 unit cells Niggli-reduced. The similarity 
			of the 3 projections is indicative of the exact or nearly
			exact rhombohedral symmetry.}
		\label{fig2}
	\end{figure}

\begin{figure}
	\includegraphics[width=1.0\textwidth]{5_21}
	\caption{The PLA2 unit cells, Niggli-reduced, and five copies were perturbed 10\%
		orthogonally to \SVI{} vector. }
	\label{fig5}	
\end{figure}

\begin{figure}
	\includegraphics[width=1.0\textwidth]{19_52}
	\label{19_57}
	\caption{The PLA2 unit cells, Niggli-reduced, and five copies were perturbed 10\%
		orthogonally to \SVI{} vector, followed by Niggli reduction.}
\end{figure}

	\begin{figure}
		\includegraphics[width=1.0\textwidth]{3_03}
		\caption{The PLA2 unit cells Delone-reduced. }
		\label{fig3}
	\end{figure}

	\begin{figure}
		\includegraphics[width=1.0\textwidth]{5_38}
		\label{5_38}
		\caption{The PLA2 cells Delone reduced and
		5 copies perturbed by 10\%.}
	\end{figure}

	\begin{figure}
		\includegraphics[width=1.0\textwidth]{6_53}
		\label{6_53}
		\caption{The same cells as plotted in Figure \ref{5_38} 
			and Delone reduced.}
	\end{figure}

\begin{figure}
	\includegraphics[width=1.0\textwidth]{7_14}
	\label{7_14}
		\caption{The same cells as plotted in Figure \ref{5_38} 
	and Delone reduced, and then each multiplied by the 24
	\SVI{} reflections.}
\end{figure}

>>>>

\documentclass[preprint]{iucr}              % DO NOT DELETE THIS LINE
\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
%\usepackage{bm}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{booktabs}
%\usepackage{calligra}
\usepackage{array}
\DeclareMathAlphabet{\mathcalligra}{T1}{calligra}{m}{n}
\def\mathbi#1{\textbf{\em #1}}
\numberwithin{equation}{section}
%\DeclareMathSymbol{\Gamma}{\mathalpha}{letters}{''00}
%\DeclareMathSymbol{\Lambda}{\mathalpha}{letters}{''03}
%\DeclareMathSymbol{\Omega}{\mathalpha}{letters}{''0A}
%\DeclareMathAlphabet{\mathitbf}{OML}{cmm}{b}{it}
\hyphenation{Niggli}
\def\mathbi#1{\textbf{\em #1}}
%\numberwithin{equation}{section}
%\DeclareMathSymbol{\Gamma}{\mathalpha}{letters}{''00}
%\DeclareMathSymbol{\Lambda}{\mathalpha}{letters}{''03}
%\DeclareMathSymbol{\Omega}{\mathalpha}{letters}{''0A}
%\DeclareMathAlphabet{\mathitbf}{OML}{cmm}{b}{it}
\usepackage{color}
\usepackage{ulem}
\usepackage{url}
\usepackage{yfonts}
%\usepackage{xr-hyper}
%\usepackage[draft]{hyperref}
%\usepackage{bibentry}
\newcommand{\SVI}[0]{$\bf{S^{6}}$}
\newcommand{\GVI}[0]{$\bf{G^{6}}$}
\newcommand{\CIII}[0]{$\bf{C^{3}}$}
\newcommand{\DVII}[0]{$\bf{D^{7}}$}
\newcommand{\VVII}[0]{$\bf{V^{7}}$}

@@

%-------------------------------------------------------------------------
% Information about journal to which submitted
%-------------------------------------------------------------------------
\journalcode{J}              % Indicate the journal to which submitted
%   A - Acta Crystallographica Section A
%   B - Acta Crystallographica Section B
%   C - Acta Crystallographica Section C
%   D - Acta Crystallographica Section D
%   E - Acta Crystallographica Section E
%   F - Acta Crystallographica Section F
%   J - Journal of Applied Crystallography
%   M - IUCrJ
%   S - Journal of Synchrotron Radiation
\makeatletter
\font\dummyft@=dummy \relax
\makeatother

@@

	% The title of the paper. Use \shorttitle to indicate an abbreviated title
	% for use in running heads (you will need to uncomment it).
	{\LARGE \emph{\today}} \\
	\title{A``Periodic Table'' for Bravais Lattices}
	%\title{Note on the transformation of three-space basis vectors to  corresponding matrix for Delaunay scalars}
	\shorttitle{A grid summarizing the Bravais lattice types according to Delaunay (Delone)}

@@

	\aff[a]{9515 NE 137th St, Kirkland, WA, 98034-1820 \country{USA}}
	\aff[b]{c/o NSLS-II, Brookhaven National Laboratory, Upton, NY, 11973 \country{USA}}

@@

	\keyword{lattice}
	\keyword{reduction}
	\keyword{Delone}
	\keyword{Selling}
	\keyword{Delaunay}

@@

	\begin{synopsis}

		A tabular arrangement of the Bravais lattice types as
		defined by Delaunay is described.

	\end{synopsis}
	\newcommand{\si}[0]{$s_1$}
	\newcommand{\sii}[0]{$s_2$}
	\newcommand{\siii}[0]{$s_3$}
	\newcommand{\siv}[0]{$s_4$}
	\newcommand{\sv}[0]{$s_5$}
	\newcommand{\svi}[0]{$s_6$}
	\newcommand{\Svec} [0] {\{\si, \sii, \siii, \siv, \sv, \svi \}}
	\newcommand{\SvecA} [0] {\{-\si, -\si+\sii, \si+\siii, \si+\sv, \si+\siv, \si+\svi \}}

@@

	\begin{abstract}

		A convenient tabular display of the Bravais lattices types
		as defined by Delaunay (using Selling reduction) is described.

@@

	\section{stuff for larry to do}
	supplementary material -- should there be a PDF file?/????

	\section{Introduction}
	\citeasnoun{kahr2023periodic} suggested a``Periodic table of the space groups''. In the
	same spirit, a ``Periodic Table'' of the Bravais lattice types is presented here.
	The table presents a systematic arrangement of the types, according to suggestions
	by \citeasnoun{Delaunay1932}. Like the periodic table of the elements, a brief amount
	of useful information is included.

\section{The space \SVI{}}

@@

	~\citeasnoun{andrews2019b} chose to 
	represent the Selling scalars in the space \SVI{},
	\Svec{} (defined in the order above), 
	as a way to create a metric space
	for the measurement of the distance between lattices.

\section{The table}

	\citeasnoun{Delaunay1932} and \citeasnoun{Delone1975} published a table of the Bravais types
	as distinguished by Selling (Delaunay) reduction.\cite{andrews2023measuring} published
	a somewhat modified version of Delaunay's table. The table presented here
	is arranged with the lattice family types as columns and the Federov polyhedron
	types as rows.
			\begin{figure}
		\includegraphics[width=10cm]{HorizontalDeloneGrid}
		\label{grid}
		\caption{The  Bravais lattice types.}
	\end{figure}

\section{The labeling of the columns (lattice families)}

	The columns are labeled according to the lattice family: cubic,
	tetragonal, rhombohedral, orthorhombic, monoclinic, and
	triclinic (anorthic). The labels for each reflect the names:
	\textbf{C}, \textbf{T}, \textbf{R}, \textbf{O},\textbf{ M}, \textbf{A}, \textbf{H} (rather than the choices of Delaunay: \textbf{K}, \textbf{Q},
	\textbf{R},\textbf{ O},\textbf{ M}, \textbf{T}, \textbf{H}, respectively).

	A further change from the model of Delaunay is the splitting 
	of orthorhombic and monoclinic into a pair of columns each
	($\bf{O_A}$, $\bf{O_B}$, $\bf{M_A}$, $\bf{M_B}$). This change 
	improved the display of these types. Delaunay had merged these
	disparate types into a single cell when their Federov
	types were identical. Thus, column $\bf{O}$ of \citeasnoun{Delaunay1932}
	has become columns $\bf{O_A}$ and $\bf{O_B}$, etc.

\section{The key cell (PQRSTU)}

The International Tables for x-ray crystallography \cite{vonLaue1952}
used the convention of labeling Selling parameters in the Bravais
tetrahedron, as \textbf{P}, \textbf{Q}, \textbf{R}, \textbf{S}, \textbf{T}, \textbf{U}, corresponding to the \SVI{} scalars
\Svec, respectively. For reference,
the figure has been included in the upper left cell of the table.
		\begin{figure}
		\includegraphics[width=5cm]{PQRSTU}
		\label{PQRSTU}
		\caption{The key cell with the labels used in the
		International Tables}
	\end{figure}

\section{The rows (Federov types)}

	\citeasnoun{fedorov1885elements} determined that there are only
	five kinds of parallelohedra that can be translated without rotations in 3-dimensional Euclidean space to fill space.

	Except for rows 3 and 4, each row has a unique number of zero
	parameters. For rows 3 and 4, the difference is whether the two
	zeros are on opposite sides of the tetrahedron or on 
	adjacent sides.

\section{The cells}

		In the table described here, the conditions for each Delone
	type are described using \SVI{}. For example, the cell for
	primitive cubic lattices is $\bf{C5}$.

	\begin{figure}
		\includegraphics[width=5cm]{C5}
		\label{C5}
		\caption{The cell for the primitive cubic Bravais lattice type.}
	\end{figure}

	 $\bf{C5}$ is the label for the position (row and column, cubic and Federov type 5 reduced cell) within
	 the table. The stylized tetrahedron shows which Selling
	 parameters are zero or equal to each other. ``cP'' is the
	 standard display for primitive cubic.

	The \SVI{} characteristic for that type is (000 rrr);

	\noindent
	\si{} = $b \cdot c$ = 0 ($\alpha$ = 90$^{\circ}$)\\
	\sii{} = $a \cdot c$ = 0 ($\beta$ = 90$^{\circ}$)\\
	\siii{} = $a \cdot b$ = 0 ($\gamma$ = 90$^{\circ}$)\\
	\siv{} = \sv{} = \svi{}

\section{Dirichlet cells}

	At the right end of the table, a few examples of Dirichlet
	cells for each row.

	\cite{fedorov1885elements}

\section{The unused/vacant cells}

	Prominent in the table are cells shown in gray. They have no
	corresponding Bravais type. For instance, the cell that would be
	$\bf{H1}$ would be a hexagonal cell with no zero parameters
	(this is no 90$^{\circ}$ angles); 
	clearly impossible.

\section{Other changes from Delaunay}

	Within the table there are five cells that have a yellow
	background. Although Delaunay identified these among the 24
	Bravais types, they are special in the sense that they
	do no possess the full dimensionality of their lattice
	family. Most likely, they are only of minor crystallographic 
	interest.

	Cells $\bf{A1}$ and $\bf{A2}$ are triclinic unit cells with 
	one or two 90$^{\circ}$ angles. Respectively, they have only five
	or four free parameters.

	$\bf{O3}$ is a 2-D manifold that is the boundary between
	 $\bf{O2}$ and $\bf{O1B}$.
	With a characteristic of (rs0 rs0), $\bf{O3}$ has only two
	free parameters, not the three that is normal for an 
	orthorhombic unit cell.

%.\cite{andrews2023measuring}
%The table of Delaunay (1932) describing the 24 Bravais types in S6. It has
%been redone removing the images of the Dirichlet cells, the non-reduced
%cells, and adding the ‘lattice character’, which describes the linear
%manifold of each type. The crystal family types have been renamed to
%modern usage: Q changed to T for tetrahedral, K changed to C for cubic
%and T changed to A for anorthic. Where Delone in some places included
%two types in one table cell, they have been split into two (for example:
%‘M1’ is split into ‘M1A’ and ‘M1B)’. Note that five types (O3, M3, M2B,
%A2 and A3) are not normal crystallographic types. They are boundary
%types, and they have fewer free parameters than the generic type requires.
%For instance, O3 (character: rs0 rs0) has only two free parameters (r and
%s), whereas an ordinary orthorhombic type requires three variables.

%\begin{figure}
%	\includegraphics[width=1.0\textwidth]{7_14}
%	\label{7_14}
%		\caption{The same cells as plotted in Figure \ref{5_38} 
%	and Delone reduced, and then each multiplied by the 24
%	\SVI{} reflections.}
%\end{figure}
